---
title: "Example Analysis for the NCEP Parasite Biodiversity: Community Data Analysis Exercise"
author: "Maureen Anne Williams"
date: "`r Sys.Date()`"
output: html_document
---
Set up the R Markdown file 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Input the data into R and load the packages needed (tidyverse, knitr, plyr) for the analysis. We need to make sure the data is in the correct form for the analysis by changing variable classifications as appropriate. 
```{r warning=FALSE, tidy=TRUE}
#Load the tidyverse package
library(tidyverse)
library(plyr)
library(knitr)

#Input dataset from csv
amnhfish <-
  read.csv("ParasiteBiodiversity_DATA.csv", #import data from CSV
           header = T,           #first row is the column names
           stringsAsFactors = F) #strings should not be immediate input as factors
str(amnhfish) #check the structre of the data set
amnhfish$fishing_status <-
  as.factor(amnhfish$fishing_status) #fishing status is a factor
amnhfish$fish_sp <-
  as.factor(amnhfish$fish_sp) #fish species is a factor
amnhfish$island <- as.factor(amnhfish$island) #island is a factor
amnhfish$habitat <-
  as.factor(amnhfish$habitat) #fish habitat is a factor
amnhfish$parasite <-
  as.factor(amnhfish$parasite) #parasite is a factor
amnhfish$parasite_class <-
  as.factor(amnhfish$parasite_class) #parasite class is a factor
amnhfish$transmission <-
  as.factor(amnhfish$transmission) #parasite transmission method is a factor
amnhfish$count <-
  as.numeric(amnhfish$count) #parasite count is a number, not an integer

```

Our main task is to find the means and standard deviations for each parasite within each fish species on fished and unfished islands. To do this, we will generate two summary tables of means and standard errors. The first is the overall mean for each parasite within each host and the second looks at fished and unfished islands in particular. 

```{r warning=FALSE, tidy=TRUE}
kable(                            #kable prints the table within the markdown document
  summary_table <- amnhfish %>%  #generate a summary table that is based on the data in amnhfish
    ddply(                       #ddply allows us to quickly summarize the data
      c("fish_sp", "parasite"),  #these are the groupings we are getting the mean for
      summarize,                 #summarize will summarize the data   
      N    = length(count),      #N is the number of individuals sampled
      mean = mean (count, na.rm = T), #mean generates the mean number of parasites, na.rm allows us to ignore any totally missing cells while still counting the zeros
      sd   = sd(count, na.rm = T), #sd generates the standard deviation of the number of parasites, na.rm allows us to ignore any totally missing cells while still counting the zeros 
      se   = sd / sqrt(N) #se generates the standard error of the number of parasites, na.rm allows us to ignore any totally missing cells while still counting the zeros
    )
)

kable(                                            #kable prints the table within the markdown document
  summary_table_fishing <- amnhfish %>%           #generate a new summary table that is based on the data in amnhfish
    ddply(                                        #ddply allows us to quickly summarize the data
      c("fish_sp", "fishing_status", "parasite"), #these are the groupings we are getting the mean for
 summarize,                 #summarize will summarize the data   
      N    = length(count),      #N is the number of individuals sampled
      mean = mean (count, na.rm = T), #mean generates the mean number of parasites, na.rm allows us to ignore any totally missing cells while still counting the zeros
      sd   = sd(count, na.rm = T), #sd generates the standard deviation of the number of parasites, na.rm allows us to ignore any totally missing cells while still counting the zeros 
      se   = sd / sqrt(N) #se generates the standard error of the number of parasites, na.rm allows us to ignore any totally missing cells while still counting the zeros
    )
)
                   
```


Now we want to generate a plot of this data to visually analyze the patterns. 
```{r}
initial_plot <-         #Generate the plot
  ggplot(summary_table_fishing,     #Use ggplot 2 and the summary table we generated above
         aes(     #set the aesthetics within ggplot2
           x = parasite, #our x axis are the parasites
           y = mean,     #our y axis is the mean of the abundance
           fill = fishing_status, #make separate bars based on fishing status 
           width = 0.75 #make the bars a little thinner so they're easier to see
         )) +
  geom_bar(stat = "identity", #make the chart a bar chart
           position = position_dodge(), #don't stack the fished and unfished data, put them side by side
           color = "black") + #outline the bar in black
  facet_wrap( ~ fish_sp, scales = "free") + #make a different panel for each host species and don't require all parasites to be represented for each fish
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), #set the standard error as the error bar
                width = .2, #make the error bars not too wide
                position = position_dodge(0.9)) + #make sure error bars are on the correct bars
  theme(           #make it slightly prettier 
    axis.text = element_text(size = 12), #set axis text size 
    axis.title = element_text(size = 14, face = "bold"), #set axis title text size 
    legend.title = element_text(size = 12), #set legend title text size 
    legend.text = element_text(size = 12), #set legend text size 
    axis.text.x = element_text(angle = 65, hjust = 1), #set axis text to be angled so it is legible and keep it underneath the axis 
    legend.position = "right", #legend is present and on the right
    panel.grid.major = element_blank(), #remove internal grids
    panel.grid.minor = element_blank(), #remove internal grids
    panel.background = element_blank(), #remove internal background color
    axis.line = element_line(colour = "black") #axis lines are black
  )
initial_plot #view the plot
```

